### 2015-11-12 ch02-qlogms/excitation-design.tex
Import QLogMS TIM paper
Some TODOs in QLogMS
emove old bib defs
Import SYSID2012 paper
Figure ID cycle + some intro exp design
Gebrol in intro/conclusie qLogMS

\chapter{Design of excitation signals}
\researchBasedOn{Geerardyn2012IMTC, Geerardyn2013TIM, Larsson2012SYSID}

@@

>>>>

\chapter{Design of excitation signals}
\def\thisDir{ch02-qlogms}
\researchBasedOn{Geerardyn2012IMTC, Geerardyn2013TIM, Larsson2012SYSID}

\section{Introduction}
\label{sec:excitation:intro}
Experiment design is the first step in the system identification loop, as illustrated in \figref{fig:intro:identification-cycle}.
Consequently, the design that is chosen influences every further step in the identification process and limit the model quality that can be attained.
In many cases

\subsection{Generic excitation signals}
\label{sec:excitation:intro:generic}

\subsection{Optimal excitation signals}
\label{sec:excitation:intro:optimal}
Since generic excitation signals are not specifically geared towards the particular system under test, such signals tend to waste signal power.

, such signals essentially don't use all the available signal power to reduce the uncertainty in the estimate.

\subsection{Robust excitation signals}
\label{sec:excitation:intro:robust}
As illustrated above, optimal experiment design places the user in a catch-22 situation.

\subsection{Approach in this chapter}
\label{sec:excitation:intro:approach}
In this chapter, the goal is to construct excitation signals that bridge the properties of generic and robust excitation signals to alleviate the chicken-and-the-egg problem that optimal strategies typically incur.
In particular, a generic excitation signal will be used but its properties are tuned in such a way that it

\input{ch02-qlogms/logMS_IMTC.tex}

\input{ch02-qlogms/sysid.tex}

@@

\section{Conclusion}
\label{sec:excitation:conclusion}
In this chapter we have illustrated that quasi-logarithmic multisine excitations are a well-suited to measure transfer functions over broad frequency ranges with a constant relative uncertainty of the present resonances.

The suggested amplitude compensation methods allow the user to measure a parametric LTI model for systems in a wide frequency band with a constant uncertainty.
The variance of the estimates at lower frequencies is improved in comparison to the uncompensated case at the cost of a slightly increased (yet constant) variance at higher frequencies for excitation signals with an equal power content.
In practical applications, this slight increase is likely to be negligible.
The suggested compensation works well as long as the considered systems are sufficiently damped with respect to the density of the excited frequency lines.

We have also shown an effective way to choose the frequency spacing of a (quasi)-logarithmically spaced multisine for an equidamped set of systems.
The presented method allows to select a maximal value for the frequency ratio $\alpha = 1 + \damping$ for which the performance will not suffer from the unavoidable misalignment of the system with respect to the excited frequencies.
In practice we advise to use an excitation signal with approximately four excited lines within the $3\unit{dB}$ bandwidth of each second-order subsystem.

\TODO{omrekentabel verschillende grootheden: $F$, $F/\mathrm{dec}$, $F/\mathrm{oct}$, $\alpha$, $\delta f$}

\begin{subappendices}
  \section{Justification for considering only poles}

\end{subappendices}

### 2015-11-12 ch02-qlogms/logMS_IMTC.tex
Gebrol in intro/conclusie qLogMS

\TODO{reshape figures}
\TODO{rename labels}
% \begin{abstract}
% The logarithmic distribution of spectral lines in excitation signals is widely used to measure the transfer functions of dynamic systems over a wide frequency band, covering several decades.
% Periodic signals are very popular in many advanced dynamic signal analyzers.
% Generating multitone periodic signals requires an equidistant frequency grid which conflicts with the logarithmic distribution, especially at the low frequencies.
% In this paper we offer an elegant solution to get around this problem using an improved choice for the amplitude spectrum of the multitone.
% We also offer a simple way to tune the frequency spacing of such a signal to allow for robust identification of the system under test.
% \end{abstract}

@@

  Such a system can be decomposed into a series of first and second order systems by the partial fraction expansion.
  To study the behavior of one sub-system, we inspect a generic second order system $G\left( s \right)$.
  In continuous time such a system can be described by a transfer function of the form
  \begin{equation}
    G\left(s \right) = \frac{\dc}
                            {             \frac{s^2}
                                               {\wn^2} 
                             + 2 \damping \frac{s}
                                               {\wn}
                             + 1
                            }
  \text{.}
  \label{eq:secondOrderSystemCT}
  \end{equation}
\TODO{This is actually wrong, rather, it should be $\frac{as-b}{cs^2 + ds + 1}$. However, the difference in information is rather small around the actual resonance. See script in NnF folder.}
\TODO{write out in partial fraction expansion, use script to show that in practice, only a few parameters matter. This will be well-approximated by considering only the second order system of the paper.}

@@

  Consequently, the complete multisine has a period of
  \begin{equation}
    T_{u(t)}  = \lcm \set{T_1, \ldots, T_F}
      = \lcm \set{f_1^{-1}, \ldots, f_F^{-1}}
    \text{.}
  \end{equation}
  \begin{definition}[Least common (integer) multiple] \label{def:lcm-real}
  The least common multiple of the set of values $T_k \in \RealNumbers$ for $k \in \set{1, \ldots, F}$ is  
  \begin{equation}
    \lcm \set{T_1, \ldots, T_F} \isdef
      \min \set{T \given{\forall k \in \set{1,\ldots,F}, \exists n_k \in \NaturalNumbers: n_k T_k = T }}
   \end{equation}
   if such a minimum exists.
   \end{definition}
  \begin{example}
   Consider a generalized multisine where $f_1 = 1 \unit{Hz}$ and $f_2 = \tfrac{2}{3} \unit{Hz}$.
   The corresponding periods are $T_1 = f_1^{-1} = 1 \unit{s}$ and $T_2 = f_2^{-1} = 1.5 \unit{s}$.
   The overall period of the multisine is $T = n_1 T_1 = n_2 T_2 = 3 \unit{s}$, i.e. $n_1 = 3$ and $n_2=2$.
  \end{example}

@@

  \begin{property}
   Each frequency band $B = \ClosedInterval{f, \kappa f}$ with $\kappa \in \PositiveRealNumbers \without \set{0,1}$ that is completely within the range of the excited logarithmic grid of frequencies $\set{f_1, \ldots, f_F}$ contains a constant number of frequency lines
   \begin{equation}
     F_{B}(f) \isdef
     \floor{\frac{\log (\kappa f) - \log( f )}{\log \alpha}}
         = \floor{\log_{\alpha} \kappa}
         \text{.}
   \end{equation}
   \end{property}
  \begin{example}
    For the frequency decade band $\decade(f) = \ClosedInterval{f, 10 f}$, a logarithmic frequency grid contains $F_{\decade} = \floor{\log_{\alpha} 10}$ excited lines.
  \end{example}

@@

  For $1/f$ noise (`pink noise'), the power spectral density $S(f)$ is proportional to $1/f$.
  We can again approximate such a signal by means of a quasi-logarithmic grid multisine.
  To do so, we excite a quasi-logarithmic frequency grid $\set{f_1,\ldots,f_k,\ldots,f_F}$ and determine the corresponding amplitude spectrum $A_k$ such that each excited line carries the same power as the corresponding frequency band does for pink noise.

@@

\TODO{zin loopt niet goed}
  The covariance matrix $\C_{G\left(\omega_{\interest}\right)}$ of the parametric transfer function, which is pretty hard because the common lack of intuition about this model given the parameters, can be shown \citep{Pintelon2012, Gallager2008, matrixcookbook} to be approximately equal to
  \begin{align}
    \C_{G\left( \omega_{\interest}\right) } \approx
      \InnerQuadraticForm[\HT]
                         {\jac{\interest}}
                         {\FisherInformation_{\theta}^{-1}}
     \label{eq:varPeak}
  \end{align}
  with 
  \begin{equation}
    \FisherInformation_{\theta} =
    2
    \real{
          \InnerQuadraticForm[\HT]
                              {\jac{\excited}}
                              {\C_{X}^{-1}}
          }
    \label{eq:FisherInfo}
  \end{equation}
  the Fisher information matrix \citep{Pintelon2012}.

@@

  Let's select a logarithmic frequency grid with an infinite number of grid lines
  \begin{equation}
    \vec{\omega} = \set{\alpha^k \omega_0 \given{k \in \ZN}}
    \label{eq:freqGridLog}
    \text{.}
  \end{equation}
  Consider a second grid, defined as a multiple of this grid, $m\vec{\omega} = \set{m \cdot \alpha^k \omega_0 \given{k \in \ZN}}$; both grids share the same frequency ratio $\alpha$.

@@

\section{Conclusion} \label{sec:conclusion}
The suggested compensation method allows the user to measure a parametric LTI model for systems in a wide frequency band with a constant uncertainty.
The variance of the estimates at lower frequencies is improved in comparison to the uncompensated case at the cost of a slightly increased (yet constant) variance at higher frequencies for excitation signals with an equal power content.
In practical applications, this slight increase is likely to be negligible.
The suggested compensation works well as long as the considered systems are sufficiently damped with respect to the density of the excited frequency lines.

We have also shown an effective way to choose the frequency spacing of a (quasi)-logarithmically spaced multisine for an equidamped set of systems.
The presented method allows to select a maximal value for the frequency factor $\alpha = 1 + \damping$ for which the performance will not suffer from the unavoidable misalignment of the system with respect to the excited frequencies.
In practice we advise to use an excitation signal with approximately four excited lines within the $3\unit{dB}$ bandwidth of each second-order subsystem.

\TODO{omrekentabel verschillende grootheden: $F$, $F/\mathrm{dec}$, $F/\mathrm{oct}$, $\alpha$, $\delta f$}

>>>>

\TODO{reshape figures}
\TODO{rename labels}

@@

  Such a system can be decomposed into a series of first and second order systems by the partial fraction expansion.

  To study the behavior of one sub-system, we inspect a generic second order system $G\left( s \right)$.
  In continuous time such a system can be described by a transfer function of the form
  \begin{equation}
    G\left(s \right) = \frac{\dc}
                            {             \frac{s^2}
                                               {\wn^2} 
                             + 2 \damping \frac{s}
                                               {\wn}
                             + 1
                            }
  \text{.}
  \label{eq:secondOrderSystemCT}
  \end{equation}
\TODO{This is actually wrong, rather, it should be $\frac{as-b}{cs^2 + ds + 1}$. However, the difference in information is rather small around the actual resonance. See script in NnF folder.}
\TODO{write out in partial fraction expansion, use script to show that in practice, only a few parameters matter. This will be well-approximated by considering only the second order system of the paper.}

@@

  Consequently, the complete multisine has a period $T_{u(t)}$ that is the common multiple of the period of all its components:
  \begin{equation}
    T_{u(t)}  = \lcm \set{T_1, \ldots, T_F}
      = \lcm \set{f_1^{-1}, \ldots, f_F^{-1}}
    \text{.}
  \end{equation}
  \begin{definition}[Least common (integer) multiple] \label{def:lcm-real}
  The least common multiple of the set of values $T_k \in \RealNumbers$ for $k \in \set{1, \ldots, F}$ is  
  \begin{equation}
    \lcm \set{T_1, \ldots, T_F} \isdef
      \min \set{T \given{\forall k \in \set{1,\ldots,F}, \exists n_k \in \NaturalNumbers: n_k T_k = T }}
   \end{equation}
   if such a minimum exists.
   \end{definition}
  \begin{example}
   Consider a generalized multisine where $f_1 = 1 \unit{Hz}$ and $f_2 = \tfrac{2}{3} \unit{Hz}$.
   The corresponding periods are $T_1 = f_1^{-1} = 1 \unit{s}$ and $T_2 = f_2^{-1} = 1.5 \unit{s}$.
   The overall period of the multisine is $T = n_1 T_1 = n_2 T_2 = 3 \unit{s}$, i.e. $n_1 = 3$ and $n_2=2$.
  \end{example}

@@

  \begin{property}
   Each frequency band $B = \ClosedInterval{f, \kappa f}$ with $\kappa \in \PositiveRealNumbers \without \set{0,1}$ that is completely within the range of the excited logarithmic grid of frequencies $\set{f_1, \ldots, f_F}$ contains a constant number of frequency lines
   \begin{equation}
     F_{B}(f) \isdef
     \floor{\frac{\log (\kappa f) - \log( f )}{\log \alpha}}
         = \floor{\log_{\alpha} \kappa}
         \text{.}
   \end{equation}
   \end{property}
  \begin{example}
    For the frequency decade band $\decade(f) \isdef \ClosedInterval{f, 10 f}$, a logarithmic frequency grid contains $F_{\decade} = \floor{\log_{\alpha} 10}$ excited lines.
  \end{example}

@@

  For $1/f$ noise (`pink noise'), the power spectral density $S(f)$ is proportional to $1/f$.
  We can again approximate such a signal by means of a quasi-logarithmic grid multisine.
  To do so, we construct a quasi-logarithmic frequency grid $\set{f_1,\ldots,f_k,\ldots,f_F}$ and determine the corresponding amplitude spectrum $A_k$ such that each excited line carries the same power as the corresponding frequency band does for pink noise.

@@

\TODO{zin loopt niet goed}
  The covariance matrix $\C_{G\left(\omega_{\interest}\right)}$ of the parametric transfer function $G(\theta,\omega)$ can be shown \citep{Pintelon2012, Gallager2008, matrixcookbook} to be approximately equal to
  \begin{align}
    \C_{G\left( \omega_{\interest}\right) } \approx
      \InnerQuadraticForm[\HT]
                         {\jac{\interest}}
                         {\FisherInformation_{\theta}^{-1}}
     \label{eq:varPeak}
  \end{align}
  with 
  \begin{equation}
    \FisherInformation_{\theta} =
    2
    \real{
          \InnerQuadraticForm[\HT]
                              {\jac{\excited}}
                              {\C_{X}^{-1}}
          }
    \label{eq:FisherInfo}
  \end{equation}
  the Fisher information matrix \citep{Pintelon2012}.

@@

  Consider a logarithmic frequency grid with an infinite number of grid lines
  \begin{equation}
    \vec{\omega} = \set{\alpha^k \omega_0 \given{k \in \ZN}}
    \label{eq:freqGridLog}
    \text{.}
  \end{equation}

  Consider a second grid, defined as a multiple of this grid, $m\vec{\omega} = \set{m \cdot \alpha^k \omega_0 \given{k \in \ZN}}$; both grids share the same frequency ratio $\alpha$.

@@
